---
title: "CDF Estimation"
author: "Michael Hornstein"
date: "11/20/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
```

### Confidence Bands 

Suppose $F(x)$ is a cdf, and $\widehat{F}_n(x)$ is an estimate of $F$ based on sample of size $n$.  Suppose someone has proven a result of the form 
\begin{equation} \label{confBandBound}
  P\left( \sup_x \left| F(x) - \widehat{F}_n(x) \right| > \varepsilon \right) \leq g(\varepsilon, n),
\end{equation}

Geometrically, the event 
\begin{equation} \label{supError}
\mathcal{E} := \left\{ \sup_x \left| F(x) - \widehat{F}_n(x) \right| \leq \varepsilon \right\}
\end{equation}
corresponds to the estimated function $\widehat{F}_n$ being within a band of width $\varepsilon$ above or below the function $F$.  Inequality \eqref{confBandBound} bounds the probability that $\mathcal{E}$ fails to occur.

Suppose $g(\varepsilon, n)$ is a decreasing function of $n$, so that as the sample size increases, the probability that the error exceeds a fixed threshold decreases.  Furthermore, note that $g(\varepsilon, n)$ is a decreasing function of $\varepsilon$, because as the threshold increases (i.e. gets more permissive), the probability that the error exceeds the threshold decreases.

Inequality \eqref{confBandBound} uses a fixed threshold $\varepsilon$ that does not depend on $n$.  But we could allow the threshold to depend on $n$ by applying the inequality using a threshold that depends on $n$, as follows: 
$$
P\left( \sup_x \left| F(x) - \widehat{F}_n(x) \right| > \varepsilon(n) \right) \leq g( \varepsilon(n), n).
$$
In particular, for a fixed $\alpha \in [0, 1]$ (typically $\alpha = 0.05$), let's solve the equation $g(\varepsilon, n) = \alpha$ for $\varepsilon$ as a function of $n$, denoting the solution as $\varepsilon_\alpha(n)$.  Then 
$$
P\left( \sup_x \left| F(x) - \widehat{F}_n(x) \right| > \varepsilon_\alpha(n) \right) \leq g( \varepsilon_\alpha(n), n) = \alpha.
$$
This inequality states that with probability at least $1 - \alpha$, the function $\widehat{F}_n$ remains within a band of $F$ of width plus/minus $\varepsilon_\alpha(n)$, i.e. 
$$ 
  \lVert F - \widehat{F}_n \rVert_\infty \leq \varepsilon_\alpha(n) \qquad \text{with probability at least $1 - \alpha$},
$$ 
where $\lVert f(x) \rVert_\infty := \sup_x |f(x)|$ is the $\ell_\infty$ norm.  

The key point is that $g(\varepsilon, n)$ in \eqref{confBandBound} does not depend on the true distribution $F$, so we end up with a bound on the error in estimating $F$ that holds with the same probability regardless of the true unknown $F$.  This is what makes it possible to form the confidence band.  

### Dvoretzky-Kiefer-Wolfowitz inequality and Confidence Bands

By the Dvoretzky-Kiefer-Wolfowitz inequality, 
\begin{equation} \label{DKW}
P\left( \sup_x \left | F(x) - \widehat{F}_n(x) \right | > \varepsilon \right) \leq 2 \exp( - 2n \varepsilon^2).
\end{equation}

Note that the right-hand side of \eqref{DKW} is a decreasing function of $n$ and $\varepsilon$.  The larger the sample size, the smaller the probaiblity that the error exceeds $\varepsilon$.  Also, the larger $\varepsilon$, the smaller the probability that the error exceeds $\varepsilon$ (because a larger $\varepsilon$ is more permissive).  

As discussed in Chapter 7 of Wasserman's All of Statistics, we can use this result to form a confidence band around the empirical cdf.

For a fixed sample size $n$, we can find the value of $\varepsilon$ for which the probability bound in \eqref{DKW} is equal to $\alpha = 0.05$.  To do so, we solve the following equation for $\varepsilon$: 
$$
2\exp( - 2n \varepsilon^2) = \alpha.
$$, 
The solution is 
$$
\varepsilon = \sqrt{ \frac{\log(\alpha / 2)}{-2n} } = \sqrt{ \frac{\log(2 / \alpha)}{2n} }.
$$
It follows that 
$$
P\left( \sup_x \left | F(x) - \widehat{F}_n(x) \right | > \sqrt{ \frac{\log(2 / \alpha)}{2n} } \right) \leq \alpha.
$$



See the link https://ggplot2.tidyverse.org/reference/stat_ecdf.html.

```{r}
n <- 1e2
x.values <- seq(from=-4, to=4, length.out=n) %>% sort
df <- data.frame(z.observed = rnorm(n), 
                 x.values, 
                 z.pdf = pnorm(x.values))
ggplot(df) + stat_ecdf(aes(z.observed)) + 
  geom_line(aes(x.values, z.pdf), col="red")
```

